### 2022-05-10 paper/localpsf.tex
update low rank update appendix based on Tucker's comments

>>>>

\section{Recycling Krylov information via low rank updates}
\label{sec:recycle_krylov}

In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems of the form
\begin{equation}
\label{eq:Ax_equals_b}
\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)},
\end{equation}
where each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. If $\mathbf{B}^{(k)}=\mathbf{A}$, then our H-matrix approximation $\mathbf{A}_H^{-1}$ may be used to precondition this linear system. If $\mathbf{B}^{(k)}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}^{(k)}$. For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ is the Hessian of the objective function in the optimization problem, and $\mathbf{f}^{(k)}$ is the negative gradient of the objective function. The Hessian and the gradient are evaluated at the $k$th Newton iterate, so both $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ change as the Newton iterations proceed. The Hessian is typically the sum of a data misfit term which is only available matrix-free, and a regularization term which is available directly as a sparse matrix. The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian. Let 
\begin{equation*}
\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
\end{equation*}
denote the H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method. The iterative method for solving the $k$th linear system will typically involve, as a sub-procedure, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$. Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$. In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}^{(k)}$ while solving the current linear system may be summarized by the equation:
\begin{equation*}
\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
\end{equation*}
For convenience, define 
\begin{equation*}
\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}, \qquad \text{and} \qquad \boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}.
\end{equation*}
One of the following three formulas may be used to form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$:
\begin{align}
	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T \label{eq:dfp_update}
\end{align}
In all three cases, direct calculation shows that we have $\mathbf{P}^{(k)}\mathbf{Y}=\mathbf{X}$ after the update, which implies $\left(\mathbf{P}^{(k)}\right)^{-1} \mathbf{X} = \mathbf{B}^{(k)}\mathbf{X}$. Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$k$ updates of $\mathbf{P}^{(k)}$, while formula \eqref{eq:dfp_update} performs a rank-$2k$ update. Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric. Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite. Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite. Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$k$'' update formula. Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell. For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}. In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.

%to update $\mathbf{B}^{(k)} \approx \left(\mathbf{A}^{(k)}\right)^{-1}$ denote the H-matrix preconditioner used for the $k$th linear system, and set $\mathbf{B}^{(1)}:= \mathbf{A}_H^{-1}$ denote the H-matrix approximation used to precondition the first linear system. Rather than re-build the H-matrix preconditioner for each subsequent linear system, one may instead form a preconditioner, $\mathbf{B}^{(k+1)}$ update, by \emph{recycling} the information generated by the iterative method used to solve the current system.
%
%In practice it is often necessary to solve a sequence of linear systems of the form 
%\begin{equation*}
%	\mathbf{A} \mathbf{x} = \mathbf{b}
%\end{equation*}
%where either $\mathbf{A}$, or $\mathbf{b}$, or both, are slowly varying. 
%
%\begin{equation*}
%	\mathbf{B}\mathbf{X} = \mathbf{Y}
%\end{equation*}
%
%
%
%\begin{equation*}
%\mathbf{Q} := \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T
%\end{equation*}
%
%\begin{equation}
%	\mathbf{B}_{k+1} = \left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right)\mathbf{B}_k\left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right) + \mathbf{Y}\boldsymbol{\Theta}\mathbf{Y}^T
%\end{equation}
